Ab Initio Effective One-Electron Potential Operators. II. Applications for Exchange-Repulsion Energy in Effective Fragment Potentials

### 2019-12-30 papers/3_oep_technique.Pauli/submit.tex
Submitted versions of papers added

\input{header.tex}
\begin{document}
%\preprint{AIP/123-OEP}

@@

The exchange\hyp{}repulsion (EXR) interaction energy 
usually describes the non\hyp{}electrostatic repulsion
between two isolated and unperturbed 
wavefunctions in an interacting
molecular 
complex.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975,Hayes.Stone.MolPhys.1984,Jeziorski.Moszynski.Szalewicz.ChemRev.1994,Jensen.Gordon.MolPhys.1996,Hesselmann.Jansen.Schutz.JCP.2005,Mandado.Hermida-Ramon.JCTC.2011}
In the
second generation of the
effective fragment potentials (EFP2) method\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
the EXR energy is an important component of the total interaction energy.
It is
derived from the intermolecular perturbation theory
of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
up to the second order with respect to the wavefunction overlap
assuming Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} (HF) wavefunctions,
and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{equation} \label{e:exr-efp2}
 E^{\rm Ex-Rep} \approx
 E^{\rm Ex} +
 E^{\rm Rep}(S^{-1}) + 
 E^{\rm Rep}(S^{-2}) \;.
\end{equation}
%
In the above equation,
$E^{\rm Ex}$ is the exchange energy,\cite{Jensen.JCP.1996}
%
\begin{equation} \label{e:exc-efp2}
 E^{\rm Ex} \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} 
 \sqrt{\frac{-2\ln{\vert S_{ij}\vert}}{\pi}} \frac{S^2_{ij}}{r_{ij}} \;,
\end{equation}
%
whereas the first\hyp{} and second\hyp{}order repulsion terms
are accordingly~\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{multline} \label{e:rep-efp2.s1}
 E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} \\
 \times
 \left\{
 \sum_{k\in A}^{\rm Locc} F_{ik}^A S_{kj} 
+\sum_{l\in B}^{\rm Locc} F_{jl}^A S_{li} 
 -2T_{ij}
 \right\}
\end{multline}
%
and 
%
\begin{multline} \label{e:rep-efp2.s2}
 E^{\rm Rep}(S^{-2}) \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij}^2 
 \\ \times 
 \Bigg\{
 \sum_{x\in A}^{\rm At}  \frac{-Z_x}{r_{xb}}
+\sum_{y\in B}^{\rm At}  \frac{-Z_y}{r_{ya}} 
+\sum_{k\in A}^{\rm Locc} \frac{   2}{r_{jk}}
+\sum_{l\in B}^{\rm Locc} \frac{   2}{r_{il}}
-r_{ij}^{-1}
 \Bigg\} \;.
\end{multline}
%
The indices $i$, $j$, $k$ and $l$ label the localized occupied orbitals (LMO's, denoted by `Locc')
located at their charge centroids ${\bf r}_{i(j)}$,
whereas $x$ and $y$ label atomic nuclei (denoted by `At') with atomic numbers 
$Z_{x(y)}$ located at ${\bf r}_{x(y)}$. The relative distances are
defined by $r_{uw}=\vert {\bf r}_u-{\bf r}_w\vert$.
$S_{ij}$ and $T_{ij}$ are the overlap and kinetic\hyp{}energy
integrals, respectively, whereas $F_{ik}^X$ are the Fock matrix elements
of unperturbed monomer $X$, all defined in LMO basis.
It is known that evaluating Eq.~\eqref{e:exr-efp2}
is remarkably efficient and gives usually very accurate
estimates of the reference EXR energies according to the symmetry 
adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT).
The computational cost to evaluate
the exchange and second\hyp{}order terms
is much smaller than that of the first\hyp{}order term,
which requires evaluation of not only the one\hyp{}electron overlap,
but also the kinetic
energy integrals in a space of occupied molecular orbital (MO) basis.

@@

\begin{multline} \label{e:rep.murrell-etal.S1.oep-v}

@@

     = -\sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} \\

>>>>

\documentclass[aip,jcp,amsmath,amssymb,preprint,floatfix]{revtex4-1}
\input{header.tex}
\begin{document}

@@

The exchange\hyp{}repulsion (EXR) interaction energy 
usually describes the non\hyp{}electrostatic repulsion
between two isolated and unperturbed 
wavefunctions in an interacting
molecular 
complex.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975,Hayes.Stone.MolPhys.1984,Jeziorski.Moszynski.Szalewicz.ChemRev.1994,Jensen.Gordon.MolPhys.1996,Hesselmann.Jansen.Schutz.JCP.2005,Mandado.Hermida-Ramon.JCTC.2011}
In the
second generation of the
effective fragment potentials (EFP2) method\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
the EXR energy is an important component of the total interaction energy.
It is
derived from the intermolecular perturbation theory
of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
up to the second order with respect to the wavefunction overlap
assuming Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} (HF) wavefunctions,
and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{equation} \label{e:exr-efp2}
 E^{\rm Ex-Rep} \approx
 E^{\rm Ex} +
 E^{\rm Rep}(S^{-1}) + 
 E^{\rm Rep}(S^{-2}) \;.
\end{equation}
%
In the above equation,
$E^{\rm Ex}$ is the exchange energy,\cite{Jensen.JCP.1996}
%
\begin{equation} \label{e:exc-efp2}
 E^{\rm Ex} \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} 
 \sqrt{\frac{-2\ln{\vert S_{ij}\vert}}{\pi}} \frac{S^2_{ij}}{r_{ij}} \;,
\end{equation}
%
whereas the first\hyp{} and second\hyp{}order repulsion terms
are accordingly~\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{equation} \label{e:rep-efp2.s1}
 E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} 
% \times
 \left\{
 \sum_{k\in A}^{\rm Locc} F_{ik}^A S_{kj} 
+\sum_{l\in B}^{\rm Locc} F_{jl}^A S_{li} 
 -2T_{ij}
 \right\}
\end{equation}
%
and 
%
\begin{equation} \label{e:rep-efp2.s2}
 E^{\rm Rep}(S^{-2}) \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij}^2 
 %\\ \times 
 \Bigg\{
 \sum_{x\in A}^{\rm At}  \frac{-Z_x}{r_{xb}}
+\sum_{y\in B}^{\rm At}  \frac{-Z_y}{r_{ya}} 
+\sum_{k\in A}^{\rm Locc} \frac{   2}{r_{jk}}
+\sum_{l\in B}^{\rm Locc} \frac{   2}{r_{il}}
-r_{ij}^{-1}
 \Bigg\} \;.
\end{equation}
%
The indices $i$, $j$, $k$ and $l$ label the localized occupied orbitals (LMO's, denoted by `Locc')
located at their charge centroids ${\bf r}_{i(j)}$,
whereas $x$ and $y$ label atomic nuclei (denoted by `At') with atomic numbers 
$Z_{x(y)}$ located at ${\bf r}_{x(y)}$. The relative distances are
defined by $r_{uw}=\vert {\bf r}_u-{\bf r}_w\vert$.
$S_{ij}$ and $T_{ij}$ are the overlap and kinetic\hyp{}energy
integrals, respectively, whereas $F_{ik}^X$ are the Fock matrix elements
of unperturbed monomer $X$, all defined in LMO basis.
It is known that evaluating Eq.~\eqref{e:exr-efp2}
is remarkably efficient and gives usually very accurate
estimates of the reference EXR energies according to the symmetry 
adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT).
The computational cost to evaluate
the exchange and second\hyp{}order terms
is much smaller than that of the first\hyp{}order term,
which requires evaluation of not only the one\hyp{}electron overlap,
but also the kinetic
energy integrals in a space of occupied molecular orbital (MO) basis.

@@

\begin{equation} \label{e:rep.murrell-etal.S1.oep-v}

@@

     = -\sum_{x\in X}^{\rm At} W_{\zeta i}^{(x)} %\\
